### 2016-06-27 ch03-lrm/nonparametric.tex
Small corrections Ch03

\paragraph*{Outline}
\secref{sec:theory} introduces theory of estimating \glspl{FRF} using local modeling techniques.
In \secref{sec:biascalc}, the bias of the \gls{LRM} is derived.
In \secref{sec:nparam:simulations} the performance of the different modeling methods are compared by means of simulations.
In \secref{sec:nparam:orderSelection}, some options for model order selection are discussed.
In \secref{sec:nonparametric:truncation}, an approach is developed to further reduce the variance of the \gls{LPM} estimate by truncation of the impulse response.
Finally, conclusions are presented in \secref{sec:conclusion}.

@@

Consider a discrete-time generalized output-error \gls{LTI} set-up, excited by the input signal $u[n]$.
For an infinite data length, such a system can be described in the time domain by
\begin{equation}
  y[n] = \ezbrace{\true{G}(q) u[n]}{\true{y}[n]} + \ezbrace{\true{H}(q) \true{e}[n]}{v[n]} \text{ with } n \in \IntegerNumbers
  \label{eq:output-error-TD-infinite}
\end{equation}
where $v[n]$ is filtered white noise, $v[n] = \true{H}(q) \true{e}[n]$ where $q^{-1}$ is the backward shift operator, i.e. $q^ {-1}x[n] = x[n-1]$.
The transfer functions $\true{G}$ and $\true{H}$ are rational functions that are causal and stable.

@@

By applying the \gls{DFT}~\eqref{eq:DFT} to both sides of \eqref{eq:output-error-TD-finite}, one obtains its frequency-domain counterpart:
\begin{align}
    Y \left( \omega_k \right) 
    & = \true{G} \left( \omega_k \right) U \left( \omega_k \right) 
      + \true{H} \left( \omega_k \right) E \left( \omega_k \right)
      + T \left( \omega_k \right)\\
      &= \true{G} \left( \omega_k \right) U \left( \omega_k \right)  + V(\omega_k) + T(\omega_k)
      \text{.}
  \label{eq:output-error-FD}
\end{align}

@@

@@

\begin{remark}
It should be noted that \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T} are essentially isomorphic to a continuous-time transfer function model with complex coefficients.
The discrete-time counterparts (i.e. substituting $d^i \to \exp (ji d \Ts)$ in the expressions), have been tried as well.
However, these preliminary experiments did not yield estimates that were numerically reliable.
This is in correspondence with the remark from \citet{Pintelon2006BJ1}.
\end{remark}

@@

\begin{remark}
Since it is well-known that the Vandemonde structure in the design matrix \eqref{eq:nparam:designMatrix} leads to numerical ill-conditioning for higher model complexities, additional measures are taken to improve numerical conditioning.
To improve numerical conditioning~\citep{Pintelon2005} of the estimation problem, we substitute $d = \dw{r}{k}$ in equations \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T}
\begin{equation}
\dw{r}{k} \isdef \frac{\omega_{k+r} - \omega_k}{\Delta \omega_k}
\label{eq:freqScaling}
\end{equation}
where
\begin{equation}
  \Delta \omega_k \isdef
  \max
  \Set{
    \abs{ \omega_k - \omega_j} |  \vphantom{\frac{a}{b}}  \omega_j \in \LocalWindow[k]
  }
\end{equation}
such that $\abs{\dw{r}{k}} \leq 1$ when $r \in \LocalShifts{k}$.
\end{remark}

@@

\subsection{The \glsentrydesc{LPM}}
\label{sec:nparam:LPM}
The \gls{LPM} is method that predates the \gls{LRM} and has been devised by~\citet{Schoukens2006LPM}.
The \gls{LPM} approximates both the \gls{FRF} and transient contribution by means of complex-valued local polynomials in the frequency domain.

@@

\begin{remark}
The trivial setting for local `modeling' where a window with a total width of $1\unit{bin}$ and no transient estimation ($\order{C} = -1$), actually corresponds to the simple \gls{ETFE}~\citep{Broersen1995,Stenman2000ASETFE,Stenman2001ASFRF} when no further smoothing or windowing is applied:
\begin{equation}
  \estimated{G}_{\ETFE}(\omega_k) \isdef \frac{Y(\omega_k)}{U(\omega_k)}
  \label{eq:nparam:ETFE}
\end{equation}
As such, $\ETFE \equiv \lrm{0,0,0,-1} \equiv \lpm{0,0,-1}$.
\end{remark}

@@

To determine the bias, an approach similar to \citep[Appendix A]{Guillaume1995} is followed.
Let us denote the expected value of the cost function over all measurements $\LocalVector{Z} = \mat{\LocalMatrix{Y}, \LocalMatrix{U}}$ as
\begin{align}
  \ELSCost{\theta}
              & \isdef 
                   \E[\LocalVector{Z}]%
                         {    
                                \LSCost{\LocalVector{\theta}, \LocalVector{Z}}} \\
              & = \ELSCost[0]{\theta} + \ELSCost[n]{\theta} \\
    \ELSCost[0]{\theta} & \isdef 
            \frac{1}{\nWind} 
                \sum_{\dW} 
                     \left| 
                             \LocalModel{A}(\dW, \LocalVector{\theta}) \true{\LocalVector{Y}}
                          - \LocalModel{B}(\dW, \LocalVector{\theta}) \true{\LocalVector{U}}
                          - \LocalModel{C}(\dW, \LocalVector{\theta}) 
                    \right|^2 \\
    \ELSCost[n]{\theta} & \isdef 
              \frac{1}{\nWind} 
                     \sum_{\dW} 
                              \left| \LocalModel{A}(\dW,\LocalVector{\theta}) \right|^2 
                              \sigma_V^2
\end{align}
where $\true{\LocalVector{U}}$ and $\true{\LocalVector{Y}}$ denote the `true' (noiseless) input-output spectra in the local window.
Note that this function is a quadratic function of $\LocalVector{\theta}$ and hence it can be represented exactly by its second order Taylor series around $\true\theta$:
\begin{multline}
  \ELSCost{\theta} = \ELSCost{\true\theta} 
  + \PartialDerivative{\ELSCost{\true\theta}}{\true\theta}  \left( \theta - \true\theta \right) %\\
  + \frac{1}{2} \left( \theta - \true\theta \right)^{\T} 
  \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta} 
   \left( \theta - \true\theta \right)
   \text{.}
\end{multline}
We use the notation
\begin{equation}
  \PartialDerivative{\ELSCost{\true\theta}}{\true\theta}
  \isdef
  \evaluate{\PartialDerivative{\ELSCost(\theta)}{\theta}}{\theta = \true\theta}
\end{equation}
for the derivative of $\ELSCost{\theta}$ with respect to $\theta$, evaluated in $\true\theta$.
The minimizer $\estimated\theta$ of this function can be obtained by equating the derivative
\begin{equation}
  \PartialDerivative{\ELSCost{\estimated\theta}}
                                          {\estimated\theta} 
  = 
  \left( \PartialDerivative{\ELSCost{\true\theta}}{\true\theta} \right)^{\T} 
  + \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta} \left( \estimated\theta - \true\theta \right)
\end{equation}
to zero.
This is equivalent to 
\begin{equation}
  \left( \estimated\theta - \true\theta \right) 
  = 
  - \left(   \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta} \right)^{-1}  
     \left( \PartialDerivative{\ELSCost{\true\theta}}{\true\theta} \right)^{\T}
  \text{.}
\end{equation}
which expresses the bias $\Delta\theta = \left( \estimated\theta - \true\theta \right)$ of the estimated parameters.
Remark that 
\begin{equation}
  \PartialDerivative{\ELSCost{\true\theta}}{\true\theta}  
  = 
  \PartialDerivative{\ELSCost[0]{\theta} + \ELSCost[n]{\theta} }{\true\theta} 
  = 
  \PartialDerivative{\ELSCost[n]{\theta} }{\true\theta} 
\end{equation}
since $\ELSCost[0]{\theta}$ is minimized in $\true\theta$ by definition.
As such, the bias on the parameters can be simplified to
\begin{equation}
  \Delta\theta = 
  - \left(   \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta^2} \right)^{-1} 
     \left( \PartialDerivative{\ELSCost[n]{\true\theta}}{\true\theta} \right)^{\T}
  \text{.}
  \label{eq:nparam:bias-basic}
\end{equation}

@@

\paragraph{Contributions to the first order derivative}
The last term in \eqref{eq:nparam:bias-basic}, i.e. the first order derivative,  is given by
\begin{equation}
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta}
  = 
  \begin{bmatrix}
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta_A} &
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta_B} &
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta_C}
  \end{bmatrix}
\end{equation}
and split into the parts pertaining to respectively $\LocalModel{A}$, $\LocalModel{B}$ and, $\LocalModel{C}$.
The respective components are given by:
\begin{align}
  \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{a_i} 
     &= 
     \frac{\sigma_V^2}{\nWind}
     \sum_{\dW} \dW^i \LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}}) 
     && \forall i \in \set{1,\ldots,\order{A}}
     \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{b_i} 
   &= 0
   && \forall i \in \set{0,\ldots,\order{B}}
   \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{c_i} 
   &= 0
   &&\forall i \in \set{0,\ldots,\order{C}}
   \text{.}
\end{align}
It can be seen that only the factors pertaining to $\theta_{A}$ are present.
As this term occurs linearly in expression \eqref{eq:nparam:bias-basic}, only $\theta_A$ can incur a bias.
Consequently, only the poles can shifted by the presence of noise.
As such, the bias term $\Delta \theta$ can be written in the following sparse form:
\begin{equation}
  \Delta \theta = 
  \begin{bmatrix}
    \Delta \theta_A\\
    \Delta \theta_B\\
    \Delta \theta_C
  \end{bmatrix}
  =
  \begin{bmatrix}
    \Delta \theta_A\\
    \deemph{0}\\
    \deemph{0}
  \end{bmatrix}
  \text{.}
  \label{eq:nparam:lrm:bias:sparse}
\end{equation}

@@

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\dW,\theta)
  \approx
  \frac{\LocalModel{G}(\dW, \theta)}
           {\LocalModel{A}(\dW, \theta)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{align}
\frac{\Delta \LocalModel{G}(\dw{\omega}{}, \theta)}{\LocalModel{G}(\dw{\omega}{}, \theta)}
&\approx
\LocalModel{A}(\dw{\omega}{}, \theta)^{-1}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}_{A}
  \PartialDerivative{\ELSCost[n]{\true\theta}}{\theta_A}^{\TT}
  \\
  &=
  \frac{\sigma_V^2}{\nWind}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}_{A}
  \begin{bmatrix}
     \Sum_{\dW} \dW^1 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \deemph{\vdots}\\
     \Sum_{\dW} \dW^{\order{A}} 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \end{bmatrix}
\end{align}
where $\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}_{A}$ denotes the upper left $\order{A}\times\order{A}$ subblock of $\left(  \PartialDerivative[2]{\ELSCost{\theta}}{ \theta^2 } \right)^{-1}$ which pertains to the parameters of $\LocalModel{A}$.
From the leading factor, it can be seen that the bias on the \gls{FRF} is $\bigO{\sigma_V^2}$.
Consequently, one can expect that the bias is inherently linked to the \gls{SNR} of the output signal in the frequency domain.

@@

The input signal used is a random-phase multisine with a full linear grid.
A number of samples of the input- and output are measured, in such a way that the \gls{DFT} grid excited $\nBW$ bins within the $3\unit{dB}$ bandwidth of the system.
The different values for 
\[\nBW \in \set{2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 14, 16, 32, 40, 64, 80, 92, 128, 256}\] are tried. 
The use of $\nBW$ makes that the experiment length is normalized to take the bandwidth of the resonance into account.
This means that the obtained results generalize to systems with a different relative damping.

The \gls{FRF} is estimated using the local modeling approaches indicated in \tabref{tbl:nparam:methods}.
This table also indicate $\Delta\SNR$ which indicates the expected gain in \gls{SNR} if the error is mainly determined by the variance.
Concretely,
\begin{equation}
  \Delta\SNR \isdef \sqrt{\frac{\nth}{\nWind}}
  \label{eq:nparam:deltaSNR}
\end{equation}
accounts for the smoothing action of local modeling.

@@

\paragraph{Good signal-to-noise ratios}
The next situation we consider is when $\SNR=60\unit{dB}$ as shown in \figref{fig:nparam:comparison:hiSNR}.
Here, the \gls{LPM} remains severely biased for small datasets ($\nBW < 12$).
However, for sufficiently long experiments ($\nBW > 32$), the bias becomes negligible.
For the \gls{LRM} and \gls{LRIC}, the bias is always much smaller than the variance.
In contrast to \gls{LPM}, the \gls{LRM} already reaches a constant variance level when $\nBW > 10$.
It should also be noted that for the \gls{LRIC}, the variance is almost $10\unit{dB}$ more than the corresponding \gls{LPM} method.
It is hence not advantageous to use the \gls{LRIC} as-is.
For all methods (except the \gls{ETFE} and \gls{LRIC}), the simulations show that the \gls{RMSE} reaches the noise floor minus $\Delta \SNR$.

For the more noisy situation where $\SNR = 40 \unit{dB}$ (\figref{fig:nparam:comparison:midSNR}), the situation is qualitatively equal except the bias incurred by the \gls{LRM} is similar to the noise level.

@@

\begin{guideline}[Use the \glsentryshort{LRM} for resonant systems when $\SNR > 10 \unit{dB}$]
For resonant systems, the \gls{LRM} offers better performance than \gls{LPM} when a good \gls{SNR} is obtained in the frequency domain.
When the \gls{SNR} is poor ($\leq 10\unit{dB}$), the \gls{LPM} may perform a slightly better than \gls{LRM}.
The \gls{LRIC} is unsuitable due to a high variance, but has a much smaller bias error than the non-iterative solutions.
\end{guideline}

\begin{guideline}[Use a second-order \glsentryshort{LRM} for resonant systems]
For resonant systems, the \gls{LRM} offers better performance than \gls{LPM} when a good \gls{SNR} is obtained in the frequency domain.
In the assumption that difference resonances are well-separated in the frequency, a second-order \gls{LRM} (e.g. $\lrm{7,2,2,2}$) is a reliable method to estimate the \gls{FRF}.
\end{guideline}

@@

\subsection{Combining Different Local Model Orders}
In the previous section, different methods were introduced to select the local model order.
This begs the question whether it is advantageous to use the same local model complexity over all frequencies or the model order should be kept globally.
The answer to this, depends on the ultimate purpose of the \gls{FRF}.

@@

For the high \gls{SNR} case, it can be seen in \figref{fig:nparam:PRESS} that the \gls{LPM} is heavily biased and hence results in a large overestimate of the noise variance near the resonance peak.
As a result, the \gls{PRESS} in the same region is also elevated for the \gls{LPM}.
Outside the resonance peak, the \gls{LPM} and \gls{LRM} estimate very similar noise variances.
The \gls{PRESS} the different methods are in agreement with the observations regarding the \gls{RMS} error from the Monte Carlo simulations in \figref{fig:nparam:comparison:midSNR}: all tested \gls{LRM} perform similarly (with a slight advantage for the first-order method), the \gls{LPM} produces much larger errors near the resonance.
Nevertheless, this figure illustrates that for the \gls{LRM}, the error is quite flat over the frequency.
Note also that the \gls{PRESS} suffers from a quite high variance over the frequency, which is a well-known drawback~\citep{Arlot2010}.

>>>>

\paragraph*{Outline}
\secref{sec:theory} introduces theory of estimating \glspl{FRF} using local modeling techniques.
In \secref{sec:biascalc}, the bias of the \gls{LRM} is derived.
In \secref{sec:nparam:simulations} the performance of the different modeling methods is compared by means of simulations.
In \secref{sec:nparam:orderSelection}, some options for model order selection are discussed.
In \secref{sec:nonparametric:truncation}, an approach is developed to further reduce the variance of the \gls{LPM} estimate by truncation of the impulse response.
Finally, conclusions are presented in \secref{sec:conclusion}.

@@

Consider a discrete-time generalized output-error \gls{LTI} set-up, excited by the input signal $u[n]$.
For an infinite data length, such a system can be described in the time domain by
\begin{equation}
  y[n] = \ezbrace{\true{G}(q) u[n]}{\true{y}[n]} + \ezbrace{\true{H}(q) \true{e}[n]}{v[n]} \text{ with } n \in \IntegerNumbers
  \label{eq:output-error-TD-infinite}
\end{equation}
where $v[n]$ is filtered white noise, $v[n] = \true{H}(q) \true{e}[n]$ where $q^{-1}$ is the backward shift operator, i.e. $q^ {-1}x[n] = x[n-1]$.
The transfer functions $\true{G}$ and $\true{H}$ are rational functions.

@@

By applying the \gls{DFT}~\eqref{eq:DFT} to both sides of \eqref{eq:output-error-TD-finite}, one obtains its frequency-domain counterpart:
\begin{align}
    Y \left( \omega_k \right) 
    & = \true{G} \left( \omega_k \right) U \left( \omega_k \right) 
      + \true{H} \left( \omega_k \right) \true{E} \left( \omega_k \right)
      + T \left( \omega_k \right)\\
      &= \true{G} \left( \omega_k \right) U \left( \omega_k \right)  + V(\omega_k) + T(\omega_k)
      \text{.}
  \label{eq:output-error-FD}
\end{align}

@@

\begin{remark}
The starting values for the \gls{LRIC} are obtained by means of the \gls{LRM}, the details of which are explained in \secref{sec:nparam:LRM}.
\end{remark}

@@

\begin{remark}
It should be noted that \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T} are essentially isomorphic to a continuous-time transfer function model with complex coefficients.
The discrete-time counterparts (i.e. substituting $d^i \to \exp (ji d \Ts)$ in the expressions), have been tried as well.
However, these preliminary experiments did not yield estimates that were numerically reliable.
This is in correspondence with the remark from \citet[Section 2.4]{Pintelon2006BJ1}.
\end{remark}

@@

\begin{remark}
Since it is well-known that the Vandermonde structure in the design matrix \eqref{eq:nparam:designMatrix} leads to numerical ill-conditioning for higher model complexities, additional measures are taken to improve numerical conditioning.
To improve numerical conditioning~\citep{Pintelon2005} of the estimation problem, we substitute $d = \dw{r}{k}$ in equations \eqref{eq:LRM:model:G} and \eqref{eq:LRM:model:T}
\begin{equation}
\dw{r}{k} \isdef \frac{\omega_{k+r} - \omega_k}{\Delta \omega_k}
\label{eq:freqScaling}
\end{equation}
where
\begin{equation}
  \Delta \omega_k \isdef
  \max
  \Set{
    \abs{ \omega_k - \omega_j} |  \vphantom{\frac{a}{b}}  \omega_j \in \LocalWindow[k]
  }
\end{equation}
such that $\abs{\dw{r}{k}} \leq 1$ when $r \in \LocalShifts{k}$.
\end{remark}

@@

\subsection{The \glsentrydesc{LPM}}
\label{sec:nparam:LPM}
The \gls{LPM} is a method that predates the \gls{LRM} and has been devised by~\citet{Schoukens2006LPM}.
The \gls{LPM} approximates both the \gls{FRF} and transient contribution by means of complex-valued local polynomials in the frequency domain.

@@

\begin{remark}
The trivial setting for local `modeling' where a window with a total width of $1\unit{bin}$ and no transient estimation ($\order{C} = -1$) are used, corresponds to the simple \gls{ETFE}~\citep{Broersen1995,Stenman2000ASETFE,Stenman2001ASFRF} when no further smoothing or windowing is applied:
\begin{equation}
  \estimated{G}_{\ETFE}(\omega_k) \isdef \frac{Y(\omega_k)}{U(\omega_k)}
  \label{eq:nparam:ETFE}
\end{equation}
As such, $\ETFE \equiv \lrm{0,0,0,-1} \equiv \lpm{0,0,-1}$.
\end{remark}

@@

To determine the bias, an approach similar to \citep[Appendix A]{Guillaume1995} is followed.

Let us denote the expected value of the cost function \eqref{eq:nparam:LRM:costFunc} over all measurements $\LocalVector{Z} = \mat{\LocalMatrix{Y}, \LocalMatrix{U}}$ as
\begin{align}
  \ELSCost{\theta}
              & \isdef 
                   \E[\LocalVector{Z}]%
                         {    
                                \LocalVector{L}_{\LRM} (\LocalVector{\theta}, \LocalVector{Z}}) \\
              & = \ELSCost[0]{\theta} + \ELSCost[n]{\theta} \\
    \ELSCost[0]{\theta} & \isdef 
            \frac{1}{\nWind} 
                \sum_{\dW} 
                     \left| 
                             \LocalModel{A}(\dW, \LocalVector{\theta}) \true{\LocalVector{Y}}
                          - \LocalModel{B}(\dW, \LocalVector{\theta}) \true{\LocalVector{U}}
                          - \LocalModel{C}(\dW, \LocalVector{\theta}) 
                    \right|^2 \\
    \ELSCost[n]{\theta} & \isdef 
              \frac{1}{\nWind} 
                     \sum_{\dW} 
                              \left| \LocalModel{A}(\dW,\LocalVector{\theta}) \right|^2 
                              \sigma_V^2
\end{align}
where $\true{\LocalVector{U}}$ and $\true{\LocalVector{Y}}$ denote the `true' (noiseless) input-output spectra in the local window.
Note that this function is a quadratic function of $\LocalVector{\theta}$ and hence it can be represented exactly by its second order Taylor series around $\true\theta$:
\begin{multline}
  \ELSCost{\theta} = \ELSCost{\true\theta} 
  + \PartialDerivative{\ELSCost{\true\theta}}{\true\theta}  \left( \theta - \true\theta \right) %\\
  + \frac{1}{2} \left( \theta - \true\theta \right)^{\T} 
  \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta^2} 
   \left( \theta - \true\theta \right)
   \text{.}
\end{multline}
We use the notation
\begin{equation}
  \PartialDerivative{\ELSCost{\true\theta}}{\true\theta}
  \isdef
  \evaluate{\PartialDerivative{\ELSCost{\theta}}{\theta}}{\theta = \true\theta}
\end{equation}
for the derivative of $\ELSCost{\theta}$ with respect to $\theta$, evaluated in $\true\theta$.
The minimizer $\estimated\theta$ of this function can be obtained by equating the derivative
\begin{equation}
  \left(\PartialDerivative{\ELSCost{\estimated\theta}}
                                          {\estimated\theta} \right)^{\T}
  = 
  \left( \PartialDerivative{\ELSCost{\true\theta}}{\true\theta} \right)^{\T} 
  + \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta^2} \left( \estimated\theta - \true\theta \right)
\end{equation}
to zero.
This is equivalent to 
\begin{equation}
  \left( \estimated\theta - \true\theta \right) 
  = 
  - \left( \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta^2} \right)^{-1}  
     \left( \PartialDerivative{\ELSCost{\true\theta}}{\true\theta} \right)^{\T}
  \text{.}
\end{equation}
which expresses the bias $\Delta\theta = \left( \estimated\theta - \true\theta \right)$ of the estimated parameters.
Remark that 
\begin{equation}
  \PartialDerivative{\ELSCost{\true\theta}}{\true\theta}  
  = 
  \PartialDerivative{\ELSCost[0]{\theta} + \ELSCost[n]{\theta} }{\true\theta} 
  = 
  \PartialDerivative{\ELSCost[n]{\theta} }{\true\theta} 
\end{equation}
since $\ELSCost[0]{\theta}$ is minimized in $\true\theta$ by definition.
As such, the bias on the parameters can be simplified to
\begin{equation}
  \Delta\theta = 
  - \left(   \PartialDerivative[2]{\ELSCost{\true\theta}}{\true\theta^2} \right)^{-1} 
     \left( \PartialDerivative{\ELSCost[n]{\true\theta}}{\true\theta} \right)^{\T}
  \text{.}
  \label{eq:nparam:bias-basic}
\end{equation}

@@

\paragraph{Contributions to the first order derivative}
The last term in \eqref{eq:nparam:bias-basic}, i.e. the first order derivative,  is given by
\begin{equation}
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta}
  = 
  \begin{bmatrix}
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta_A} &
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta_B} &
  \PartialDerivative{\ELSCost[n]{\true\theta, \conj{\true\theta}}}{\theta_C}
  \end{bmatrix}
\end{equation}
and split into the parts pertaining to respectively $\LocalModel{A}$, $\LocalModel{B}$ and, $\LocalModel{C}$.
The respective components are given by:
\begin{align}
  \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{a_i} 
     &= 
     \frac{\sigma_V^2}{\nWind}
     \sum_{\dW} \dW^i \LocalModel{A}(\conj{\dW},\conj{\LocalVector{\theta}}) 
     && \forall i \in \set{1,\ldots,\order{A}}
     \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{b_i} 
   &= 0
   && \forall i \in \set{0,\ldots,\order{B}}
   \\
   \PartialDerivative{\ELSCost[n]{\theta,\conj{\theta}}}{c_i} 
   &= 0
   &&\forall i \in \set{0,\ldots,\order{C}}
   \text{.}
\end{align}
It can be seen that only the factors pertaining to $\theta_{A}$ are present.
As this term occurs linearly in expression \eqref{eq:nparam:bias-basic}, only $\theta_A$ can incur a bias.
Consequently, only the poles can be shifted by the presence of output noise.
As such, the bias term $\Delta \theta$ can be written in the following sparse form:
\begin{equation}
  \Delta \theta = 
  \begin{bmatrix}
    \Delta \theta_A\\
    \Delta \theta_B\\
    \Delta \theta_C
  \end{bmatrix}
  =
  \begin{bmatrix}
    \Delta \theta_A\\
    \deemph{0}\\
    \deemph{0}
  \end{bmatrix}
  \text{.}
  \label{eq:nparam:lrm:bias:sparse}
\end{equation}

@@

However, the sparsity of $\Delta\theta$, see \eqref{eq:nparam:lrm:bias:sparse}, can be exploited such that the bias on the \gls{FRF} can be expressed as:
\begin{equation}
  \Delta \LocalModel{G}(\dW,\theta)
  \approx
  - 
 \frac{\LocalModel{G}(\dW, \theta)}
           {\LocalModel{A}(\dW, \theta)}
  \begin{bmatrix}
    \dW^1 & \deemph{\cdots} & \dW^{\order{A}}
  \end{bmatrix}
  \Delta \theta_A
  \text{.}
\end{equation}
Equivalently, the relative bias of the \gls{FRF} can be written as
\begin{align}
\frac{\Delta \LocalModel{G}(\dw{\omega}{}, \theta)}
          {\LocalModel{G}(\dw{\omega}{}, \theta)}
&\approx
-
\LocalModel{A}(\dw{\omega}{}, \theta)^{-1}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\true\theta}}{ \theta^2 } \right)^{-1}_{A}
  \PartialDerivative{\ELSCost[n]{\true\theta}}{\true{\theta_A}}^{\TT}
  \\
  &=
  -
  \frac{\sigma_V^2}{\nWind}
\begin{bmatrix}
    \dw{\omega}{}^1 & \deemph{\cdots} & \dw{\omega}{}^{\order{A}}
  \end{bmatrix}
  \left(  \PartialDerivative[2]{\ELSCost{\true\theta}}{ \true\theta^2 } \right)^{-1}_{A}
  \begin{bmatrix}
     \Sum_{\dW} \dW^1 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\true{\LocalVector{\theta}}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \deemph{\vdots}\\
     \Sum_{\dW} \dW^{\order{A}} 
     \frac{\LocalModel{A}(\conj{\dW},\conj{\true{\LocalVector{\theta}}})}
               {\LocalModel{A}(\dw{\omega}{}, \theta)}\\
  \end{bmatrix}
\end{align}
where $\left(  \PartialDerivative[2]{\ELSCost{\true\theta}}{ \true\theta^2 } \right)^{-1}_{A}$ denotes the upper left $\order{A}\times\order{A}$ sub-block of $\left(  \PartialDerivative[2]{\ELSCost{\true\theta}}{ \true\theta^2 } \right)^{-1}$ which pertains to the parameters of $\LocalModel{A}$.
From the leading factor, it can be seen that the bias on the \gls{FRF} is $\bigO{\sigma_V^2}$.
Consequently, one can expect that the bias is inherently linked to the \gls{SNR} of the output signal in the frequency domain.

@@

The input signal used is a random-phase multisine with a full linear grid.
A number of samples of the input and output are measured, in such a way that the \gls{DFT} grid excited $\nBW$ bins within the $3\unit{dB}$ bandwidth of the system.
The different values for 
\[\nBW \in \set{2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 14, 16, 32, 40, 64, 80, 92, 128, 256}\] are tried. 
The use of $\nBW$ makes that the experiment length is normalized to take the bandwidth of the resonance into account.
This means that the obtained results generalize to systems with a different relative damping.

The \gls{FRF} is estimated using the local modeling approaches indicated in \tabref{tbl:nparam:methods}.
This table also indicate $\Delta\SNR$ which indicates the expected gain in \gls{SNR} if the error is mainly determined by the variance.
Concretely,
\begin{equation}
  \Delta\SNR \isdef \frac{\nth}{\nWind}
  \label{eq:nparam:deltaSNR}
\end{equation}
accounts for the smoothing action of local modeling.

@@

\paragraph{Good signal-to-noise ratios}
The next situation we consider is when $\SNR=60\unit{dB}$ as shown in \figref{fig:nparam:comparison:hiSNR}.
Here, the \gls{LPM} remains severely biased for small datasets ($\nBW < 12$).
However, for sufficiently long experiments ($\nBW > 32$), the bias becomes negligible.
For the \gls{LRM} and \gls{LRIC}, the bias is always much smaller than the variance.
In contrast to \gls{LPM}, the \gls{LRM} already reaches a constant variance level when $\nBW > 10$.
It should also be noted that for the \gls{LRIC}, the variance is almost $10\unit{dB}$ more than the corresponding \gls{LRM} method.
It is hence not advantageous to use the \gls{LRIC} as-is.
For all methods (except the \gls{ETFE} and \gls{LRIC}), the simulations show that the \gls{RMSE} reaches the noise floor minus $\Delta \SNR$.

For the more noisy situation where $\SNR = 40 \unit{dB}$ (\figref{fig:nparam:comparison:midSNR}), the situation is qualitatively equal except the bias incurred by the \gls{LPM} reaches the level of the noise floor for a smaller data set ($\nBW = 12$).
It can be seen that the bias of the \gls{LPM} behaves independently of the noise level ($40\unit{dB}$ versus $60\unit{dB}$), while the bias of the  \gls{LRM} clearly depends on the noise level.

@@

\begin{guideline}[Use the \glsentryshort{LRM} for resonant systems when $\SNR > 10 \unit{dB}$]
For resonant systems, the \gls{LRM} offers better performance than \gls{LPM} when a good \gls{SNR} is obtained in the frequency domain.
When the \gls{SNR} is poor ($\leq 10\unit{dB}$), the \gls{LPM} may performs slightly better than \gls{LRM}.
The \gls{LRIC} is unsuitable due to a high variance, but has a much smaller bias error than the non-iterative solutions.
\end{guideline}

\begin{guideline}[Use a second-order \glsentryshort{LRM} for resonant systems]
For resonant systems, the \gls{LRM} offers better performance than \gls{LPM} when a good \gls{SNR} is obtained in the frequency domain.
In the assumption that different resonances are well-separated in the frequency, a second-order \gls{LRM} (e.g. $\lrm{7,2,2,2}$) is a reliable method to estimate the \gls{FRF}.
\end{guideline}

@@

\subsection{Combining Different Local Model Orders}
In the previous section, different methods were introduced to select the local model order.
This begs the question whether it is advantageous to use the same local model complexity globally over all frequencies or whether the local model order can be allowed to vary for different frequency bands.
In fact, both approaches have their merits and which of both is to be preferred, depends on the ultimate purpose of the \gls{FRF}.

@@

For the high \gls{SNR} case, it can be seen in \figref{fig:nparam:PRESS} that the \gls{LPM} is heavily biased and hence results in a large overestimate of the noise variance near the resonance peak.
As a result, the \gls{PRESS} in the same region is also elevated for the \gls{LPM}.
Outside the resonance peak, the \gls{LPM} and \gls{LRM} estimate very similar noise variances.
The \gls{PRESS} of the different methods are in agreement with the observations regarding the \gls{RMS} error from the Monte Carlo simulations in \figref{fig:nparam:comparison:midSNR}: all tested \gls{LRM} perform similarly (with a slight advantage for the first-order method), the \gls{LPM} produces much larger errors near the resonance.
Nevertheless, this figure illustrates that for the \gls{LRM}, the error is quite flat over the frequency.
Note also that the \gls{PRESS} suffers from a quite high variance over the frequency, which is a well-known drawback~\citep{Arlot2010}.
